Lesson 3
========================================================

***

### What to Do First?
Notes:

***

### Pseudo-Facebook User Data
Notes:

```{r Pseudo-Facebook User Data}
facebook = read.csv("pseudo_facebook.tsv", sep = '\t')
names(facebook)
```

***

### Histogram of Users' Birthdays
Notes:

```{r Histogram of Users\' Birthdays}
library(ggplot2)

qplot(data = facebook, x= dob_day) +
  scale_x_continuous(breaks=1:31) 
```

***

#### What are some things that you notice about this histogram?
Response:

***

### Moira's Investigation
Notes:

***

### Estimating Your Audience Size
Notes:

***

#### Think about a time when you posted a specific message or shared a photo on Facebook. What was it?
Response:

#### How many of your friends do you think saw that post?
Response:

#### Think about what percent of your friends on Facebook see any posts or comments that you make in a month. What percent do you think that is?
Response:

***

### Perceived Audience Size
Notes:

***
### Faceting
Notes:

```{r Faceting}
qplot(data = facebook, x= dob_day) +
  scale_x_continuous(breaks=1:31)+ 
  facet_wrap(~dob_month, ncol=3)
```

#### Let’s take another look at our plot. What stands out to you here?
Response:

***

### Be Skeptical - Outliers and Anomalies
Notes:

***

### Moira's Outlier
Notes:
#### Which case do you think applies to Moira’s outlier?
Response:

***

### Friend Count
Notes:

#### What code would you enter to create a histogram of friend counts?

```{r Friend Count}
names(facebook)
qplot(data = facebook, x= friend_count) 
```

#### How is this plot similar to Moira's first plot?
Response:

***

### Limiting the Axes
Notes:

```{r Limiting the Axes}
qplot(data = facebook, x= friend_count, xlim = c(0, 1000)) 
#OR, you can do like so
qplot(data = facebook, x= friend_count) +
  scale_x_continuous(limits = c(0, 1000))
```

### Exploring with Bin Width
Notes:

***

### Adjusting the Bin Width
Notes:

### Friend Count
```{r Friend Count}
qplot(data = facebook, x = friend_count, binwidth = 10) +
  scale_x_continuous(limits = c(0, 1000),
                     breaks = seq(0, 1000, 50)) 

#binwidth is the width of the bar. So, from 0 to 50 there's 5 bars.

#seq is used to customize x-axis, and make it starts from 0, ends at 1000 with step of 50 
#so, it goes 0, 50, 100, 150, ... 1000

```

### Faceting Friend Count
```{r Faceting Friend Count}
# What code would you add to create a facet the histogram by gender?
# Add it to the code below.
qplot(data = facebook, x = friend_count, binwidth = 10) +
  scale_x_continuous(limits = c(0, 1000),
                     breaks = seq(0, 1000, 50)) +
  facet_wrap(~gender, ncol=2)

```

***

### Omitting NA Values
Notes:

```{r Omitting NA Values}
#NA: stands for (Not Applicable). in other words, the gender infromation for some accounts was #missing

#we can exclude any NA value by using is.na() function which returns true if the value is NA
qplot(data = subset(facebook, !is.na(gender)), x = friend_count, binwidth = 10) +
  scale_x_continuous(limits = c(0, 1000),
                     breaks = seq(0, 1000, 50)) +
  facet_wrap(~gender, ncol=2)
```

***

### Statistics 'by' Gender
Notes:

```{r Statistics \'by\' Gender}
table(facebook$gender)
by(facebook$friend_count, facebook$gender, summary)
```

#### Who on average has more friends: men or women?
Response: women

#### What's the difference between the median friend count for women and men?
Response: 22

#### Why would the median be a better measure than the mean?
Response: because these histograms are positive skewed historgrams

***

### Tenure
Notes:

```{r Tenure}
#Tenure: means days of the year that the user was on facebook

qplot(data = facebook, x= tenure, 
      color= I('#000000'), fill = I('#EFE4B0'))
```

***

#### How would you create a histogram of tenure by year?

```{r Tenure Histogram by Year}

qplot(data = facebook, x= tenure/365, binwidth = 0.25,  
      color= I('#000000'), fill = I('#EFE4B0')) +
  scale_x_continuous(limits = c(0, 7),
                     breaks = seq(0, 7, 1))
  
```

***

### Labeling Plots
Notes:

```{r Labeling Plots}
qplot(data = facebook, x= tenure/365, binwidth = 0.25,
      xlab = "Number of years using facebook",
      ylab = "Number of sample users",
      color= I('#000000'), fill = I('#EFE4B0')) +
  scale_x_continuous(limits = c(0, 7),
                     breaks = seq(0, 7, 1))
```

***

### User Ages
Notes:

```{r User Ages}
qplot(data = facebook, x= age, binwidth = 2,
      color= I('#000000'), fill = I('#EF4A21')) + 
  scale_x_continuous(limits = c(10, 90),
                     breaks = seq(10, 90, 10))
  
```

#### What do you notice?
Response:

***

### The Spread of Memes
Notes:

***

### Lada's Money Bag Meme
Notes:

***

### Transforming Data
Notes:

```{r}
library(gridExtra)
p1 = qplot(data = facebook, x= friend_count ,
            xlab = "Friend Count",
           color = I('#000000'), fill = I('#0E73EF'))

p2 = qplot(data = facebook, x= log10(friend_count + 1),
            xlab = "Log10 of Friend Count",
           color = I('#000000'), fill = I('#EF0008'))

p3 = qplot(data = facebook, x= sqrt(friend_count + 1),
            xlab = "sqrt of Friend Count",
           color = I('#000000'), fill = I('#00EF30'))

grid.arrange(p1, p2, p3)

```

***

### Add a Scaling Layer
Notes:

```{r Add a Scaling Layer}

```

***


### Frequency Polygons

```{r Frequency Polygons}
qplot(data = subset(facebook, !is.na(gender)), x = friend_count, 
      binwidth = 10) +
  scale_x_continuous(limits = c(0, 1000),
                     breaks = seq(0, 1000, 50)) +
  facet_wrap(~gender, ncol=2)

#Frequency Polygons which does the same as above:
qplot(data = subset(facebook, !is.na(gender)), x = friend_count,
      binwidth = 10, geom = "freqpoly", color= gender) +
  scale_x_continuous(limits = c(0, 1000),
                     breaks = seq(0, 1000, 50))
  
#change y-axis to be the 'proportion' not the 'count'
qplot(data = subset(facebook, !is.na(gender)), 
      x = friend_count, y = ..count../sum(..count..), 
      xlab = "Friend Count",
      ylab= "Proportion of total users",
      binwidth = 10, geom = "freqpoly", color= gender) +
  scale_x_continuous(limits = c(0, 1000),
                     breaks = seq(0, 1000, 50))
```

***

### Likes on the Web
Notes:

```{r Likes on the Web}
summary(facebook$www_likes)
qplot(data = subset(facebook, !is.na(gender)), x = www_likes) +
  scale_x_continuous(limits = c(0, 4000),
                     breaks = seq(0, 4000, 250)) +
  scale_y_continuous(limits = c(0, 3500)) +
  facet_wrap(~gender, ncol=2)

#Frequency Polygon
qplot(data = subset(facebook, !is.na(gender)), x = www_likes, 
      geom = "freqpoly", color = gender) +
  scale_x_continuous(limits = c(0, 4000),
                     breaks = seq(0, 4000, 250)) +
  scale_y_continuous(limits = c(0, 3500))

#Which gender has more likes

by(facebook$www_likes, facebook$gender, sum)
```


***

### Box Plots
Notes:

```{r Box Plots}
qplot(data = subset(facebook, !is.na(gender)), 
      x=gender, y= friend_count, geom = 'boxplot') +
  scale_y_continuous(limits = c(0,1000))

#the next code is the same as the previous, but it's better as it doesn't delete any datasets.
qplot(data = subset(facebook, !is.na(gender)), 
      x=gender, y= friend_count, geom = 'boxplot') +
  coord_cartesian(ylim= c(0, 1000))
  
```

#### Adjust the code to focus on users who have friend counts between 0 and 1000.

```{r}
qplot(data = subset(facebook, !is.na(gender)), 
      x=gender, y= friend_count, geom = 'boxplot') +
  coord_cartesian(ylim= c(0, 250))

```

***

### Box Plots, Quartiles, and Friendships
Notes:

```{r Box Plots, Quartiles, and Friendships}
qplot(data = subset(facebook, !is.na(gender)), 
      x=gender, y= friend_count, geom = 'boxplot') +
  coord_cartesian(ylim= c(0, 250))

by(facebook$friend_count, facebook$gender, summary)
```


#### On average, who initiated more friendships in our sample: men or women?
Response:
#### Write about some ways that you can verify your answer.
Response:
```{r Friend Requests by Gender}
qplot(data = subset(facebook, !is.na(gender)), 
      x=gender, y= friendships_initiated, geom = 'boxplot') +
  coord_cartesian(ylim= c(0, 130))

by(facebook$friendships_initiated, facebook$gender, summary)

```

Response:

***

### Getting Logical
Notes:

```{r Getting Logical}
#this variable is used to get the people who has ever liked on a mobile device
has_mobile <- NA
facebook$has_mobile <- ifelse(facebook$mobile_likes > 0, TRUE, FALSE)
facebook$has_mobile <- factor(facebook$has_mobile)
summary(facebook$has_mobile)

has_mobile_percentage = length(which(facebook$has_mobile == TRUE)) * 100 / length(facebook$has_mobile)

round(has_mobile_percentage, 4)

```

Response:

***

### Analyzing One Variable
Reflection:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!