Lesson 4
========================================================

***

### Scatterplots and Perceived Audience Size
Notes:

***

### Scatterplots
Notes:

```{r Scatterplots}
library(ggplot2)
facebook <- read.csv("pseudo_facebook.tsv", sep='\t')

qplot(data = facebook, x= age, y= friend_count)

```

***

#### What are some things that you notice right away?
Response:

***

### ggplot Syntax
Notes:

```{r ggplot Syntax}
#here, we are going to produce ggplot which does the same thing as qplot
ggplot(data = facebook, aes(x = age, y = friend_count)) + 
  geom_point() + 
  scale_x_discrete(limits = c(13, 90))

summary(facebook$age)

```

***

### Overplotting
Notes:

```{r Overplotting}

ggplot(data = facebook, aes(x = age, y = friend_count)) + 
  geom_point(alpha = 1/20) + 
  scale_x_discrete(limits = c(13, 90))
#alpha is used to set the transparency, so if alpha = 1/20 this means that
#it takes 20 point to be located in the same place to get one black dot.

```

#### What do you notice in the plot?
Response:

***

### Coord_trans()
Notes:

```{r Coord_trans()}
ggplot(data = facebook, aes(x = age, y = friend_count)) + 
  geom_point(alpha = 1/20) + 
  scale_x_discrete(limits = c(13, 90)) +
  coord_trans( y = "sqrt")
```


#### What do you notice?

***

### Alpha and Jitter
Notes:

```{r Alpha and Jitter}

ggplot(data = facebook, aes(x = age, y = friendships_initiated)) + 
  geom_point(alpha = 1/20) +
  scale_x_discrete(limits = c(13, 90), breaks= seq(13, 90, 5))


```

***

### Overplotting and Domain Knowledge
Notes:

***

### Conditional Means
Notes:

```{r Conditional Means}

library(dplyr)
#let's group the whole facebook dataset's age.
#group_by() funcion takes two arguments, the whole dataset and what you want to group
age_group = group_by(facebook, age)

#let's make some mathematical operation with age_group like determining mean and median
#the mean is going to be calculated and put in 'friend_count_mean' vector
#the median is going to be calculated and put in 'friend_count_median' vector
#summarise() used to create a table
#n() function is used to get the number of people in each group, and it's used only with summarise()
friend_count_by_age = summarise(age_group, 
                          friend_count_mean = mean(friend_count),
                          friend_count_median = median(friend_count),
                          nummber_of_people = n())

head(friend_count_by_age, 20)
#which gives the first 20 row from the table
```

Create your plot!

```{r Conditional Means Plot}

ggplot(data = friend_count_by_age, 
       aes(x = age, y = friend_count_mean)) +
       geom_line() 


```

***

### Overlaying Summaries with Raw Data
Notes:

```{r Overlaying Summaries with Raw Data}
#here, we are going to plot the scatter plot of friend count over the whole facebook dataset
# and overlay the graph line of mean
#we are going to choose a brighter color to be able to draw the graph line above it

p <- ggplot(data = facebook, aes(x= age, y = friend_count)) + 
          geom_point(alpha = 1/20, color = "#EFE4B0") + 
          coord_cartesian(xlim= c(13, 75), ylim = c(0, 1000))

#add a graph line which represents the mean of the whole graph. 
p <- p + geom_line(stat = "summary", fun.y = mean)

#add another graph line which represents the 10% quantile (first 10% of whole dataset friend_count).
# linetype to make the line dashed
p <- p + geom_line(stat="summary", fun.y = quantile, fun.args=list(probs=0.1), color = "#0C00EF", linetype = 2)

#add another graph line which represents the 90% quantile.
p <- p + geom_line(stat="summary", fun.y = quantile, fun.args=list(probs=0.9), color = "#0C00EF", linetype = 2)

#add another graph line which represents the median (50%).
p <- p + geom_line(stat="summary", fun.y = quantile, fun.args=list(probs=0.5), color = "#EF0014")

p

```

#### What are some of your observations of the plot?
Response:

***

### Moira: Histogram Summary and Scatterplot
See the Instructor Notes of this video to download Moira's paper on perceived audience size and to see the final plot.

Notes:

***

### Correlation
Notes:

```{r Correlation}
#here, we are going to get correlation between 'age' and 'friend_count' in 'facebook' dataset
#we can get the correlation using cor.test() function
cor.test(facebook$age, facebook$friend_count, 
         method = c("pearson"))
#if the answer is less than 0.3 or -0.3, then there's a weak relation
#if it's 0.5, then it's moderate
#if it's higher than 0.7, then it's pretty strong

##NOTE##
#the correlation doesn't depend on which parameter is in x-axis or y-axis

#we can do the same but with using with() function like so:
with(facebook, cor.test(age, friend_count, method="pearson"))

```

Look up the documentation for the cor.test function.

What's the correlation between age and friend count? Round to three decimal places.
Response:

***

### Correlation on Subsets
Notes:

```{r Correlation on Subsets}
#with() function is pretty awesome with subsetting.
#let's use the correlation with datasets below 70 years old
with(subset(facebook, age<=70), cor.test(age, friend_count, method="pearson"))


```

***

### Correlation Methods
Notes:

***

## Create Scatterplots
Notes:

```{r}
ggplot(data = facebook, aes(x= likes_received, y = www_likes_received)) + 
          geom_point()


```

***

### Strong Correlations
Notes:

```{r Strong Correlations}
#here, we are going to plot the 95% from the previous data. 
#we can do that using quantile() function
p <- ggplot(data = facebook, aes(x= likes_received, y = www_likes_received)) + 
        geom_point() +
        xlim(0, quantile(facebook$likes_received, 0.95)) + 
        ylim(0, quantile(facebook$www_likes_received, 0.95))
p
```

```{r Smoother}
#here, we are going to add the 'correlation line' which is the best line that fits through 
#these datasets. We can do that by adding a smoother like so
p<- p + geom_smooth(method='lm', color= '#E70C00')
p
#'lm' means 'linear model' which represents 'linear relation'
```


What's the correlation betwen the two variables? Include the top 5% of values for the variable in the calculation and round to 3 decimal places.

```{r Correlation Calcuation}
cor.test(facebook$likes_received, facebook$www_likes_received)       

```

Response:

***

### Moira on Correlation
Notes:

***

### More Caution with Correlation
Notes:

```{r More Caution With Correlation}
install.packages('alr3')
library(alr3)
?Mitchell
head(Mitchell)
```

Create your plot!

```{r Temp vs Month}

ggplot(data= Mitchell, aes(x= Month, y= Temp)) +
  geom_point()
```

***

### Noisy Scatterplots
a. Take a guess for the correlation coefficient for the scatterplot.

b. What is the actual correlation of the two variables?
(Round to the thousandths place)

```{r Noisy Scatterplots}
cor.test(Mitchell$Month, Mitchell$Temp, method="pearson")


```

***

### Making Sense of Data
Notes:

```{r Making Sense of Data}
#we are going to break the month parameter into a year
ggplot(data= Mitchell, aes(x= Month, y= Temp)) +
        geom_point() +
        scale_x_continuous(breaks= seq(0, 203, 12))
        


```

***

### A New Perspective

What do you notice?
Response:

Watch the solution video and check out the Instructor Notes!
Notes:

***

### Understanding Noise: Age to Age Months
Notes:

```{r Understanding Noise: Age to Age Months}

ggplot(data=friend_count_by_age, aes(x=age, y= friend_count_mean)) + 
  geom_line()

```

***

### Age with Months Means

```{r Age with Months Means}

facebook$age_in_month <- facebook$age + (12 - facebook$dob_month) / 12
head(facebook$age_in_month)

#making sure everything is ok
names(facebook)

length(facebook$age)
length(facebook$dob_month)
length(facebook$age_in_month)
```

Programming Assignment
```{r Programming Assignment}

library(dplyr)
#let's group the whole facebook dataset's age_in_month.
age_group = group_by(facebook, age_in_month)

friend_count_by_age_in_month = summarise(age_group, 
                               friend_count_mean = mean(friend_count),
                               friend_count_median = median(friend_count),
                               n = n())

head(friend_count_by_age_in_month)
#which gives the first 6 rows from the table

#make sure everything is ok about my limits
range(facebook$age_in_month)


```

***

### Noise in Conditional Means

```{r Noise in Conditional Means}
#the whole dataset of all ages
ggplot(data=friend_count_by_age_in_month, aes(x=age_in_month, y=friend_count_mean)) + 
    geom_line()

#with age less than 71 years old
ggplot(data=subset(friend_count_by_age_in_month, age_in_month<71), aes(x=age_in_month, y=friend_count_mean)) + 
    geom_line()
```

***

### Smoothing Conditional Means
Notes:

```{r Smoothing Conditional Means}
#here, we are going to know how to smooth any graph line

############## WITHOUT SMOOTHING #######################
#this is a graph line made by 'age' and 'friend_count_mean' in 'friend_count_by-age' dataset
p1<- ggplot(data=subset(friend_count_by_age, age<71), aes(x=age, y=friend_count_mean)) + 
            geom_line()

#this is a graph line made by 'age_in_month' and 'friend_count_mean' in 'friend_count_by-age_in_month' dataset
p2<- ggplot(data=subset(friend_count_by_age_in_month, age_in_month<71), aes(x=age_in_month, y=friend_count_mean)) + 
    geom_line()

library(gridExtra)
grid.arrange(p1, p2)
############## WITH SMOOTHING #######################
p1<- ggplot(data=subset(friend_count_by_age, age<71), aes(x=age, y=friend_count_mean)) + 
            geom_line() + 
            geom_smooth()

#this is a graph line made by 'age_in_month' and 'friend_count_mean' in 'friend_count_by-age_in_month' dataset
p2<- ggplot(data=subset(friend_count_by_age_in_month, age_in_month<71), aes(x=age_in_month, y=friend_count_mean)) + 
    geom_line() +
    geom_smooth()


grid.arrange(p1, p2)
```

***

### Which Plot to Choose?
Notes:

***

### Analyzing Two Variables
Reflection:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

