setwLesson 5
========================================================

### Multivariate Data
Notes:

***

### Moira Perceived Audience Size Colored by Age
Notes:

***
### Load data
Notes:

```{r Load Data}
library(ggplot2)
facebook <- read.csv("pseudo_facebook.tsv", sep='\t')
```


### Third Qualitative Variable
Notes:

```{r Third Qualitative Variable}
#draw two box plots, one for each gender versus the age
ggplot(data = subset(facebook, !is.na(gender)), aes(x = gender, y = age)) + 
       geom_boxplot()

#draw two graph lines, one for each gender versus the age and the friend_count
ggplot(data = subset(facebook, !is.na(gender)), aes(x = age, y = friend_count)) + 
       geom_line(aes(color = gender), stat = "summary", fun.y= median)
#in the last plot, we have plotted three variables into one graph
# - the gender
# - the median of friend_count
# - the age 
```

***

### New Data Frames
Notes:

```{r creating data frames}
library(dplyr)
#deleting Na gender from 'facebook' dataset
#here, you can group more than one variable like so:
age_gender_group = group_by(subset(facebook, !is.na(gender)), age, gender)


friend_count_by_age_gender =  summarise(age_gender_group, 
                              mean_friend_count = mean(friend_count),
                              median_friend_count = median(friend_count),
                              nummber_of_people = n())

head (friend_count_by_age_gender, 10)

```



### Plotting Conditional Summaries
Notes:

```{r Plotting Conditional Summaries}
ggplot(data = friend_count_by_age_gender, aes(x = age, y = median_friend_count)) + 
       geom_line(aes(color = gender))

head(friend_count_by_age_gender)

```

***

### Thinking in Ratios
Notes:

***

### Wide and Long Format
Notes:

***

### Reshaping Data
Notes:

```{r}

library(reshape2)

friend_count_by_age_gender_wide = dcast(friend_count_by_age_gender, 
                                        age ~ gender, 
                                        value.var = 'median_friend_count')

head(friend_count_by_age_gender_wide)
```


***

### Ratio Plot
Notes:

```{r Ratio Plot}
#it is a good thinking to put the ratio of two parameters in mind,
#here, we are going to plot the age with the ratio of friend_count of female over male

#plot the age with the ratio
p <- ggplot(friend_count_by_age_gender_wide, aes(x= age, y= female/male)) +
    geom_line() 

#add horizontal dashed line at y=1
p <- p + geom_hline(yintercept=1, color = "#E70C00", linetype= 2)
# The linetype parameter can take the values 0-6:
# 0 = blank, 1 = solid, 2 = dashed
# 3 = dotted, 4 = dotdash, 5 = longdash
# 6 = twodash
p  

```

***

### Third Quantitative Variable
Notes:

```{r Third Quantitative Variable}
#create 'year_joined' variable which contains the year when the user joined facebook

facebook$year_joined = floor(2014 - (facebook$tenure / 365)) 

summary(facebook$year_joined)
table(facebook$year_joined)


```

***

### Cut a Variable
Notes:

```{r Cut a Variable}
#exclude the NA for year_joined
facebook <- subset(facebook, !is.na(year_joined))
#function cut() is used to divide the whole dataset into certain groups:

####based on given values
c1 <- cut(facebook$year_joined, c(2004,2009, 2011, 2012, 2014))
table(c1)

####based on equal intervals 
c1 <- cut(facebook$year_joined, breaks = 4) #four equal intervals
table(c1)

####based on certain increment 
c1 <- cut(facebook$year_joined, c(seq(2004,2012,1 ))) #starting from 2004, ending at 2014 with 1 year as increment
table(c1)
```

***

### Plotting it All Together
Notes:

```{r Plotting it All Together}

#here, each bucket of year will have a different color
ggplot(data = facebook, aes(x= age, y= friend_count)) +
      geom_line(aes(color = c1), stat = "summary", fun.y = median)

```

***

### Plot the Grand Mean
Notes:

```{r Plot the Grand Mean}
#in the last plot, we were plotting the median
#here, we are going to plot the mean
#and also add the 'grand mean'

p <- ggplot(data = facebook, aes(x= age, y= friend_count)) +
      geom_line(aes(color = c1), stat = "summary", fun.y = mean)
p <- p + geom_line(stat = "summary", fun.y = mean, linetype = 2)

p

```

***

### Friending Rate
Notes:

```{r Friending Rate}
#here, we are going to determine the rate of friend_count which equal (friend_count/tenure)
r <- subset(facebook, facebook$tenure >=1) 
summary (r$friend_count/r$tenure)
```

***

### Friendships Initiated
Notes:

What is the median friend rate?

What is the maximum friend rate?

```{r Friendships Initiated}
#remove users which have less than 1 day tenure
facebook <- subset(facebook, tenure >=1)

#create 'year_joined' variable which contains the year when the user joined facebook
facebook$year_joined = floor(2014 - (facebook$tenure / 365)) 

#exclude the NA for year_joined
facebook <- subset(facebook, !is.na(year_joined))

#cut the 'whole days'year_joined' into certain buckets
c1 <- cut(facebook$year_joined, c(2004,2009, 2011, 2012, 2014))
table(c1)

ggplot(data = facebook, aes(x= tenure, y= friendships_initiated/tenure)) + 
  geom_line(aes(color = c1), stat="summary", fun.y = mean)


# or for short, this is the graph in one line
ggplot(data = subset(facebook, tenure >= 1), aes(x = tenure, y = friendships_initiated / tenure))        +  geom_line(aes(color = c1), stat = 'summary', fun.y = mean)



```

***

### Bias-Variance Tradeoff Revisited
Notes:

```{r Bias-Variance Tradeoff Revisited}
#this is the same plot as in the last excercise
ggplot(data = subset(facebook, tenure >= 1), aes(x = tenure, y = friendships_initiated / tenure))        +  geom_line(aes(color = c1), stat = 'summary', fun.y = mean)

#let's smooth it a little bit by rounding x-values 
ggplot(data = subset(facebook, tenure >= 1), aes(x = 7 * round(tenure / 7), 
                                                 y = friendships_initiated / tenure)) + geom_line(aes(color = c1), stat = 'summary', fun.y = mean)


#let's smooth it more 
ggplot(data = subset(facebook, tenure >= 1), aes(x = 30 * round(tenure / 30), 
                                                 y = friendships_initiated / tenure)) + geom_line(aes(color = c1), stat = 'summary', fun.y = mean)

#let's smooth it even more 
ggplot(data = subset(facebook, tenure >= 1), aes(x = 90 * round(tenure / 90), 
                                                 y = friendships_initiated / tenure)) + geom_line(aes(color = c1), stat = 'summary', fun.y = mean)

```

***
### Smoother using geom_smooth
```{r smoothing}

#this is the main graph with has a lot of noise
p<- ggplot(data = subset(facebook, tenure >= 1), 
           aes(x = tenure, y = friendships_initiated / tenure)) + 
          geom_line(aes(color = c1), stat = 'summary', fun.y = mean)

#let's add the smoother
p <- p + geom_smooth()
p


######BUT #####
#if you want to plot the smoother only, then do like this
#delete the geom_line() and add geo_smooth() with color attribute
p<- ggplot(data = subset(facebook, tenure >= 1), 
           aes(x = tenure, y = friendships_initiated / tenure)) + 
          geom_smooth(aes(color = c1))
p
```



### Sean's NFL Fan Sentiment Study
Notes:

***

### Introducing the Yogurt Data Set
Notes:
```{r Yougurt}
#read the 'yogurt' dataset
yo <- read.csv("yogurt.csv")

#get to know the data
str(yo)
head (yo)

#change the 'id' variable from 'int' to 'factor'
yo$id = factor(yo$id)

#make sure it's turned into 'factor'
str(yo)
```

***

### Histograms Revisited
Notes:

```{r Histograms Revisited}
ggplot(data = yo, aes(x= price)) + 
  geom_histogram()

```

***

### Number of Purchases
Notes:

```{r Number of Purchases}
#here, we are going to add a new column called 'all.purchases' which has the sum of all types of 
#yogurt per observation. 
#we can do that using transform() function
yo = transform(yo,  all.purchases = strawberry + blueberry + pina.colada + plain + mixed.berry)
head(yo)

#or with() function
yo$all.purchases = with (yo, strawberry + blueberry + pina.colada + plain + mixed.berry)
head(yo)

# or you can use this simple code, which will take a lot of effort
yo$all.purchases = yo$strawberry + yo$blueberry + yo$pina.colada + yo$plain + yo$mixed.berry
head(yo)
```

***

### all.purchases plot
Notes:

```{r all.purchases plot}
ggplot(data= yo, aes(x=all.purchases)) +
  geom_histogram(binwidth = 1, fill = "#FF00AE")

```

***
### Prices over Time
Notes:

```{r Prices over Time}
ggplot(data= yo, aes(x=time, y=price)) +
  geom_point() 
  

```

***

### Sampling Observations
Notes:

***

### Looking at Samples of Households

```{r Looking at Sample of Households}
#it's always a good behavior of data analysist to look at some samples 
#So, here we are going to look at 16 random samples out of the whole dataset
#let's choose random 16 samples using sample() function
set.seed(0)
test_samples <- sample(yo$id, 16)
test_samples

#let's draw these samples in one graph 
p <- ggplot(data = subset(yo, id%in% test_samples), aes(x= time, y= price)) + 
        facet_wrap(~ id) +
        geom_line()
# add a different-sized point to the graph based on all.purchases
p <- p + geom_point(aes(size=all.purchases), pch=1)
p
  
```

***

### The Limits of Cross Sectional Data
Notes:

***

### Many Variables
Notes:

***

### Scatterplot Matrix
Notes:
```{r Scatterplot Matrix}
#Scatterplot Matrix is a tool which analysists use.
#it's a tool which plots correlations between all columns in the dataset
install.packages('GGally')
library(GGally)

#here, we are going to use 'facebook' dataset but we don't need the whole columns
#we are gonna exclude the first and last
temp <- facebook[, c(2:15)]
names(temp)

#let's choose 10 random rows out of 'temp' dataset
temp_samples <- temp [sample(nrow(temp), 10),]
length(temp_samples)  #14
nrow(temp_samples)    #10
temp_samples

#so now, let's draw the scatterplot Matrix
#Note: it takes really really long time, around (7) minutes
ggpairs(temp_samples)


```


***

### Even More Variables
Notes:
```{r nci dataset}
nci <- read.table("nci.tsv")

#name the columns as numbers, we, here, have 64 columns.
colnames(nci) <- c(1:64)

#discover the dataset
names(nci)
str(nci)
head(nci)
```
***

### Heat Maps
Notes:



```{r Heat Maps}
#didn't understand the coming line 
nci.long.samp <- melt(as.matrix(nci[1:200,]))
names(nci.long.samp) <- c("gene", "case", "value")
head(nci.long.samp)

ggplot(aes(y = gene, x = case, fill = value),
  data = nci.long.samp) +
  geom_tile() +
  scale_fill_gradientn(colours = colorRampPalette(c("blue", "red"))(100))
```


***

### Analyzing Three of More Variables
Reflection:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

